Equivalence Testing and the Second Generation \emph{P}-Value

### 2019-03-10 reply_to_reviews.R1.Rmd

---
title             : "Reply to Reviews"
shorttitle        : "Reply R1"

author:
  - name          : "Dani&euml;l Lakens"
    affiliation   : "1"
    corresponding : yes
    address       : "Den Dolech 1, IPO 1.33, 5600 MB, Eindhoven, The Netherlands"
    email         : "D.Lakens@tue.nl"
  - name          : "Marie Delacre"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "Eindhoven University of Technology, Eindhoven, The Netherlands"
  - id            : "2"
    institution   : "Service of Analysis of the Data, Université Libre de Bruxelles, Belgium"

abstract: |
  We respond to all the comments on our submission that were raised by the two reviewers, as well as the comments you highlighted as editor. We would like to thank you and reviewers for suggestions for improvements, which we believe made our main argument much clearer, focussed, and formally correct. We feel we were able to accomodate all suggestions, and we hope you think our submission is ready to be accepted for publication in Meta-Psychology.

@@

Dear editor,

Thank you for your comments on our submitted manuscript. We have incorporated your three small suggestions:

1. We moved the first sentence on page 3 into the next paragraph (well spotted).
2. The code for the Shiny app was on GitHub, and the GitHub repository is linked to the OSF project page, but we have added a component to the OSF page and also directly added the Shiny App code there to make it easier to find: https://osf.io/7f8uq/
3. We now immediately refer to Figure 4 in the beginning of the paragraph in which the Figure is described.

We will respond to the points raised by both reviewers in turn. Our points below detail the main issues you highlighted in your editorial letter, namely 1) that we limit ourselves to a comparison of SGPV and TOST, and do not require researchers to perform both an equivalence test and a NHST (and therefore we did not add these curves to the Figures), 2) we rewrote the frequency claims, 3) we provide more detail about confidence intervals around correlations, and 4) removed all statements about the SGPV being a descriptive statistic.

@@

# Reply to Reviewer 1

1. We have changed the part of the sentence "Our conclusions about the relationship between TOST *p*-values and SGPV in this article are not dependent upon any specific example" into "Our conclusions about the relationship between TOST *p*-values and SGPV hold for second generation *p*-values calculated from confidence intervals, and assuming data is sampled from a bivariate normal distribution". We mainly intended to refer to our conclusion that: "When confidence intervals are symmetric we can think of the SGPV as a straight line that is directly related to the *p*-value from an equivalence test for three values" namely *p* = 0.5 for TOST when the SGPV will be 0.5, and the *p*-values for TOST of 0.025 and 0.975 (corresponding to the largest observed mean where the SGPV is 0 and the smallest observed mean where the SGPV is 1). We have now made it clearer that when we talk about 'correspondence between TOST *p*-values and the SGPV', we mean that there is a direct relation between three points. We have added the explanation: "When we discuss the relationship between the *p*-values from TOST and the SGPV, we focus on their correspondence at three values, namely where the TOST *p* = 0.025 and SGPV is 1, where the TOST *p* = 0.5 and SGPV = 0.5, and where the TOST *p* = 0.975 and SGPV = 1. These three values are important for the SGPV because they indicate the values at which the SGPV indicates the data should be interpreted as compatible with the null hypothesis (SGPV = 1), or with the alternative hypothesis (SGPV = 0), or when the data are strictly inconclusive (SGPV = 0.5)."

2. We appreciate correcting our use of the term 'unrelated' and teaching us the word 'bijective'. We used the word 'unrelated' somewhat naievely, looking at the Figures and wanting to express that across the three identified situations, a different TOST *p*-value does not imply a different SGPV - the SGPV across these ranges of means is aways the same (either 0, 0.5, or 1) while the TOST *p*-value differs across this range of means. We now just directly say what we originally meant, and no longer use the word 'unrelated'. For example, "There are three situations where *p*-values for TOST differentiate between observed results, while the SGPV does not differentiate" and "Where the SGPV is 1 as long as the confidence interval falls completely within the equivalence bounds, the *p*-value for the TOST continues to differentiate between results as a function of how far the confidence interval lies within the equivalence bounds" and similarly "A third situation in which the SGPV remains stable across a range of observed effects, while the TOST *p*-value continues to differentiate, is whenever the CI is wider than the equivalence range, and the CI overlaps with the upper *and* lower equivalence bound". We have removed the word 'unrelated' throughout the manuscript.

3. Please do not feel like you ever need to apologize for having formal training we lack, that allows you to point out whenever we make formally incorrect statements. This is exactly why we submit our work for peer review - to be able to improve it based on feedback from experts. We now preface the sentence with "As long as data are sampled from a bivariate normal distribution,". We follow your recommendation throughout to not generalize beyond the scenarios we study. In the second paragraph of the manuscript we now clearly state: "We limit our analysis to continuous data sampled from a bivariate normal distribution." See also point 5 below.

4. After making the change at point 2 above, we now think the sentence "The conclusion of this section on asymmetric confidence intervals is that a SGPV of 1 or 0 can still be interpreted as a p < 0:025 or p > 0:975 in an equivalence test, since the SGPV and *p*-value for the TOST procedure are always directly related at these values." should no longer be confusing. To make the sentence clearer, we have changed "these values" into "at the values *p* = 0.025 and *p* = 0.975" It should now be clear the sections are not contradictory - TOST differentiates below *p* < 0.025, when the SGPV is 1 for all *p* < 0.025.

5. We agree that is unclear whether the original statement by Blume et al that "the degree of overlap conveys how compatible the data are with the null premise" holds in general. We have now amended the sentence to read: "Although a SGPV of 1 or 0 can be directly interpreted, a SGPV between 0 and 1 is not interpretable as 'compatibility with the null hypothesis' under the assumption of a bivariate normal distribution, and the generalizability of this statement needs to be examined beyond normal bivariate distributions."

6. We have removed all our earlier statements that the SGPV only works as a descriptive statistic. We deleted the sentence "the SGPV is a descriptive statistic (unlike the *p*-value that is calculated for an equivalence test, which is an inferential statistic) that provides the proportion of overlap of the confidence interval and the equivalence range." and deleted "The novelty of the SGPV lies in its use as a descriptive statistic" from the conclusion. Indeed, we acknowledge that "Although a SGPV of 1 or 0 can be directly interpreted, a SGPV between 0 and 1 is not interpretable as “compatibility with the null hypothesis”." So the SGPV is an inferential statistic when interpreted dichotomously. We will leave our uncertainty of how specific SGPV values (e.g., SGPV = 0.3) should be interpreted to future researchers, and we do not make any statements about whether such values are descriptive or inferential.

Based on your recommendations, we have limited our conclusions to the specific cases we have examined, and noted the possibility for future research to examine a wider set of assumptions. An additional change not mentioned above is that we renamed the section on asymmetric confidence intervals to the more specific "The relation between equivalence tests and SGPV for asymmetrical confidence intervals around correlations" instead of the more general "The relation between equivalence tests and SGPV when confidence intervals are not symmetrical". We appreciate the very positive evaluation of our manuscript, but even more so the criticism which allowed us to improve the manuscript.

@@

# Reply to Reviewer 2

Before listing major points, the reviewer raises the point that we compare SGPV with *p*-values from TOST throughout our paper, and in all figures, but then at the end, in page 20, we wrote:

@@

"Equivalence tests combined with null-hypothesis significance tests also allow researchers to conclude an effect is significant and equivalent (i.e., statistically different from zero, but also too small to be considered meaningful). Thus, the SGPV is used to classify results into one of three possible categories (with the data falling inside or outside the equivalence range, or being inconclusive), while equivalence tests combined with null-hypothesis tests classify results into four possible categories."

The reviewer says that "the authors suddenly switch the premise; now the SGPV should be replaced by two *p*-values, one from an equivalence test and one from a NHST". However, we do not suggest to replace the SGPV with an equivalence test and a NHST at all. In our manuscript, we already pointed out how the SGPV is basically the same as an equivalence test and a minimum effect test (on page 7), and we have expanded this text to clarify that a minimum effects test and TOST are complementary:

"The third point where the SGPV and the *p*-value from the TOST procedure should overlap is where the 95% CI falls completely outside of the equivalence range, but one endpoint of the confidence interval is equal to the equivalence bound (see situation C in Figure 3), when the *p*-value will always be 0.975. Note that this situation is in essence a minimum-effect test [@murphy_statistical_2014]. The goal of a minimum-effect is not just to reject a difference of zero, but to reject the smallest effect size of interest (i.e., the equivalence bounds). An equivalence test and minimum effect test against the same equivalence bound are complementary, and when a TOST *p*-value is larger than 0.975, the *p*-value for the minimum effect test is smaller than 0.05 (and therefore the minimum effect test provides no additional information that can not be derived from the *p*-value from the equivalence test). The SGPV summarizes the information from an equivalence test (and the complementary minimum-effect test)."

The overlap of the TOST *p*-value at 0.975 when SGPV = 0 gives away that the equivalence test can be interpreted as a minimum effect test against the equivalence bound. When the TOST *p*-value is 0.975, the 95% CI touches either the upper or lower equivalence bound, and fall completely outside of the equivalence range. This is mathematically identical to a statistically significant minimum effect test at *p* < 0.05. We clearly should have made this more explicit, and hope the additional text clarifies the statistical relationship between TOST, a minimum effect test, and the SGPV

@@

One might wonder why then, on page 20, why we introduce the idea to perform both a null-hypothesis significance test and an equivalence test. We were not clear enough. And indeed, the sentence "This interpretation of a SGPV as allowing researchers to reject the null, reject the presence of a meaningful effect, or remaining inconclusive is very similar to the Neyman-Pearson interpretation of combining a null-hypothesis test and an equivalence test" was simply confusing. We have deleted this sentence and replaced it with "As already mentioned, this interpretation of a SGPV as is very similar to the Neyman-Pearson interpretation of an equivalence test and a minimum effect tests (which are complementary)." Furthermore, to prevent the apparent confusion about why we suggested to perform a null-hypothesis test and equivalence test: part of this confusion was our original sentence comparing the SGPV and a combination of NHST and an equivalence test which was not needed, and we understand it made it seem as if SGPV needs to be replaced by a combination of other tests. But that is not what we propose. We rewrote the paragraph:

"Some researchers might be interested in combining an equivalence tests with a null-hypothesis significance test. This allows a researcher to ask whether there is an effect that is statistically different from zero, and whether effect sizes that are considered meaningful can be rejected. Equivalence tests combined with null-hypothesis tests classify results into four possible categories, and for example allow researchers to conclude an effect is significant and equivalent (i.e., statistically different from zero, but also too small to be considered meaningful, see (Lakens et al., 2018))."

We hope this makes it clear we are not pulling a bait and switch. The SGPV is strongly related to an equivalence test (or it's complement, a minimum effect test). Describing the relation between the SGPV and TOST is our main contribution. We just wanted to mention that researchers might want to combine an equivalence test with a NHST, because in the literature, this is a question we see many psychologists ask when they use TOST.

@@

We will now discuss the major points in turn.

1. In line with the comments by reviewer 1, in the second paragraph of the manuscript we now clearly state: "We limit our analysis to continuous data sampled from a bivariate normal distribution." and  "Our conclusions about the relationship between TOST *p*-values and SGPV hold for second generation *p*-values calculated from confidence intervals, and assuming data is sampled from a bivariate normal distribution.". After all, *p*-values and confidence are directly related, but Bayesian credible intervals would not necessarily be, depending on the prior. Blume et al (2018) limit their examples to confidence intervals, but explicitly acknowledge other possibilities (such as Bayesian intervals). We look forward to future papers that illustrate the SGPV with other interval estimates and are happy to add the limitation to continuous normally distributed data in our article.

2. The reviewer notes that the correspondence only holds when SE's are small compared to the width of the equivalence range. The correspondence we refer to is not in the shape of the curve (which clearly differs across Figures 2, 6, and 9). We clearly state that: "When confidence intervals are symmetric we can think of the SGPV as a straight line that is directly related to the *p*-value from an equivalence test for three values." This correspondence is not dependent upon the SE's. It will always hold for CI based on bivariate continuous normally distributed data. The reviewer says that "The paper does not provide theoretical justification for this connection nor any proofs." We do not think proof is needed (the 'proof' is that the SGPV and the TOST both use the same confidence interval) and we believe the explanation in the paper of why this overlap must be so is very clear. A confidence interval and a *p*-value are directly related. The TOST procedure is a test using the confidence interval, and the SGPV (now that we clearly limited our discussion to a SGPV based on a confidence interval calculated from continuous normal data) compared against the same equivalence bounds will be related as we describe. But we believe our article might not have been clear enough about that we limit our discussion about the correspondence of the curves to these 3 points. Indeed, in some figures, the curves are less similar (e.g., Figure 6, 9, 13). Nevertheless, in these graphs, we still have the correspondence between these 3 points (or at least the 2 points if the CI is too wide compared to the equivalence bounds to ever observe a *p* = 0.025). To make it clearer that we talk about the overlap in these 3 critical points, we have added:

@@

"When we discuss the relationship between the *p*-values from TOST and the SGPV, we focus on their correspondence at three values, namely where the TOST *p* = 0.025 and SGPV is 1, where the TOST *p* = 0.5 and SGPV = 0.5, and where the TOST *p* = 0.975 and SGPV = 1. These three values are important for the SGPV because they indicate the values at which the SGPV indicates the data should be interpreted as compatible with the null hypothesis (SGPV = 1), or with the alternative hypothesis (SGPV = 0), or when the data are strictly inconclusive (SGPV = 0.5). These three points of overlap are indicated by the horizontal dotted lines in Figure 2."

We believe this satisfactorily addresses the point by the reviewer that "Because of this lack of discussion, the reader can easily miss that the close correspondence only occurs when the sample size is large enough that the standard error is very small relative to the width of the equivalence zone." Indeed, the correspondence between the TOST and SGPV at these 3 points should be obvious to anyone who understands the relationship between a *p*-value and a confidence interval, so we do not provide any further proof, but we hope that clarifying we talk about the correspondence of these 3 points, and not the entire curve, is sufficient to address point 2. We have also changed the figures, making the SGPV a solid line, to make the relationship more clearly visible compared to the previous dotted curve.

@@

3. We are aware of the Beger and Hsu (1996) article. But if the reviewer reads more closely, it should be clear that we never say the TOST is UMP. We say that TOST is a test procedure that aims to control error rates - our use of the term 'Neyman-Pearson perspective' refers to the classic difference between significance tests (Fisher) and hypothesis tests (Neyman-Pearson). We do not imply that any test that fixes an error rate is a NP test. Because many psychologists are not trained in the difference between a Fisherian approach and a Neyman Pearson approach to the use of *p*-values, we are simply adding a 'Neyman-Pearson' approach in the text to explain that "From a Neyman-Pearson perspective on statistical inferences the main conclusion is based on whether or not $p < \alpha$, and thus an equivalence test and SGPV can be performed by simply checking whether the confidence interval falls within the equivalence range, just as a null-hypothesis test can be performed by checking whether the confidence interval contains zero or not." We believe this is a correct use of the term 'Neyman-Pearson perspective on statistical inferences'.

4. We tried to find the passage the reviewer refers to when he writes "In light of this, the assertion without proof that classical *p*-values provide more or better assessments of the “evidence” appear out of place." but we could not. We do not use the word 'evidence' in the text (because neither the *p*-value from TOST nor the SGPV are measures of 'evidence' - we fully agree with Bume et al., 2018, in this respect). We also do not say the *p*-value is 'naturally informative'. Indeed, related to point 3 above, we tried to be consistent in interpreting the TOST procedure from a Neyman-Pearson perspective, and focus on error control. So we regrettably do not know how to incorporate this point.

@@

5. We fully agree with the reviewer on point 5 that "The maximum Type I Error rate for the SGPV over the null interval is 5%, exactly like the TOST test." Indeed, given the direct overlap between the TOST *p*-value and the SGPV at *p* = 0.025 and 0.0975 and SGPV of 1 and 0, any error rates (when concluding support for the null or alternative when the SGPV is either 1 or 0) must be identical. This is why we were so surprised when Blume et al (2018) write "They control the Type I error naturally, forcing it to zero as the sample size grows. This, in turn, offsets Type I Error inflation that results from multiple comparisons or multiple examinations of accumulating data. Findings identified by second-generation *p*-values are less likely to be false discoveries than findings identified by classical *p*-values. Consequently, second-generation *p*-values do not require ad-hoc adjustments to provide strict error control and this improves power in studies with massive multiple comparisons." Clearly, it can not both be true that the error rate of SGPV and TOST is identical, TOST requires adjusting for multiple comparisons, but the SGPV does not.

The editor asks us to better substantiate the frequency claims for the comparison of TOST and SGPV, but since the SGPV and TOST *p*-value are directly related (a TOST *p* < 0.025 = a SGPV of 1) we do not think any additional substantiation is possible, especially since the reviewer agrees that (as long as confidence intervals are used to calculate the SGPV, which we now clearly limit our conclusions to) "The maximum Type I Error rate for the SGPV over the null interval is 5%, exactly like the TOST test".

Nevertheless, we agree this discussion was a bit too extensive. Given my own fondness of error control in statistical inferences, I get worried if researchers suggest we can ignore it, especially when making multiple comparisons. Imagine people start to use the SGPV instead of the TOST procedure whenever they perform 20 tests, and then argue they do not need to correct for multiple comparisons - I think we can all agree this is undesireable, given that we are trying to fix the replicability crisis in psychology! We have reduced the two paragraphs to one hopefully clear statement about the need to control error rates, also when using the SGPV. We hope this addresses the concerns by the editor as well as the reviewer, while still making an important point in the literature about the need to control error rates (if one cares about drawing incorrect conclusions).

"@blume_second-generation_2018 claim that when using the SGPV "Adjustments for multiple comparisons are obviated" (p. 15). However, this is not correct. Given the direct relationship between TOST and SGPV highlighted in this manuscript (where a TOST *p* = 0.025 equals SGPV = 1, as long as the SGPV is calculated based on confidence intervals, and assuming data are sampled from a continuous bivariate normal distribution), not correcting for multiple comparisons will inflate the probability of concluding the absence of a meaningful effect based on the SGPV in exactly the same way as it will for equivalence tests. Whenever statistical tests are interpreted as support for a hypothesis (e.g., SPGV = 0 or SGPV = 1), it is possible do so erroneously, and if researchers want to control error rates, they need to correct for multiple comparisons."

@@

2. This text was deleted (see major point 5 above)

### 2019-03-10 manuscript.R1.Rmd
R1 final

Second generation *p*-values (SGPV) were recently proposed to as a statistic that represents ‘the proportion of data-supported hypotheses that are also null hypotheses’ [@blume_second-generation_2018]. 
The researcher specifies an equivalence range around a null hypothesis of values that are considered practically equivalent to the null hypothesis. 
The SGPV measures the degree to which a set of data-supported parameter values falls within the interval null hypothesis. 
If the estimation interval falls completely within the equivalence range, the SGPV is 1. 
If the confidence interval falls completely outside of the equivalence range, the SGPV is 0. Otherwise the SGPV is a value between 0 and 1 that expresses the overlap of data-supported hypotheses and the equivalence range. 
When calculating the SGPV the set of data-supported parameter values can be represented by a confidence interval (CI), although one could also choose to use credible intervals or Likelihood support intervals (SI). 
When a confidence interval is used, the SGPV and equivalence tests such as the Two One-Sided Tests (TOST) procedure [@lakens_equivalence_2017;@meyners_equivalence_2012;@quertemont_how_2011;@schuirmann_comparison_1987] appear to have close ties, because both tests compare a confidence interval against an equivalence range.
Here, we aim to examine the similarities and differences between the TOST procedure and the SGPV. 
We limit our analysis to continuous data sampled from a bivariate normal distribution.

@@

The second generation *p*-value (SGPV) is calculated as:
$$
  p _ { \delta } = \frac { \left| I \cap H _ { 0 } \right| } { | I | } \times \max \left\{ \frac { | I | } { 2 \left| H _ { 0 } \right| } , 1 \right\}
$$
where I is the interval based on the data (e.g., a 95% confidence interval) and H~0~ is the equivalence range. The first term of this formula implies that the second generation *p*-value is the width of the confidence interval that overlaps with the equivalence range, divided by the total width of the confidence interval. The second term is a 'small sample correction' (which will be discussed later) that comes into play whenever the confidence interval is more than twice as wide as the equivalence range.
To examine the relation between the TOST *p*-value and the SGPV we can calculate both statistics across a range of observed effect sizes. In Figure \ref{fig:TOSTSGPV1} *p*-values are plotted for the TOST procedure and the SGPV. The statistics are calculated for hypothetical one-sample *t*-tests for observed means ranging from 140 to 150 (on the x-axis). The equivalence range is set to 145 ± 2 (i.e., an equivalence range from 143 to 147), the observed standard deviation is assumed to be 2, and the sample size is 30. For example, for the left-most point in Figure \ref{fig:TOSTSGPV1} the SGPV and the TOST *p*-value is calculated for a hypothetical study with a sample size of 30, an observed standard deviation of 2, and an observed mean of 140, where the *p*-value for the equivalence test is 1, and the SGPV is 0. Our conclusions about the relationship between TOST *p*-values and SGPV hold for an SGPV calculated from confidence intervals, and assuming continuous normally distributed data. Readers can explore the relationship between TOST *p*-values and SGPV for themselves in an online Shiny app: http://shiny.ieis.tue.nl/TOST_vs_SGPV/.

@@

  res <- TOSTone.raw(m = m, 
                     mu = mu,
                     sd = sd, 
                     n = n, 
                     low_eqbound = low_eqbound, 
                     high_eqbound = high_eqbound, 
                     alpha = alpha,
                     plot = FALSE,
                     verbose = FALSE
  )
  t <- (m - mu)/(sd/sqrt(n))
  t_list[count] <- t
  sgpv_list[count] <- p_delta(mu+res$LL_CI_TTEST, mu+res$UL_CI_TTEST, mu+low_eqbound, mu+high_eqbound)
  p_tost_list[count] <- max(res$TOST_p1, res$TOST_p2)
  p_list[count] <- 2 * pt(-abs(t), df = n-1)
}
```
```{r, TOSTSGPV1, fig.cap="Comparison of *p*-values from TOST (black line) and SGPV (dotted grey line) across a range of observed sample means (x-axis) tested against a mean of 145 in a one-sample *t*-test with a sample size of 30 and a standard deviation of 2."}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 1001),
     yaxt = "n",
     xaxt = "n",
     ylab = "P-value for TOST and SGPV",
     xlab = "Observed Mean")
axis(1, at = seq(0,1000,100), labels = seq(140,150,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)
lines(sgpv_list, type="l", col = "darkgrey", lwd = 3, lty = 3)
lines(p_tost_list, lwd = 3)
```

The SGPV treats the equivalence range as the null-hypothesis, while the TOST procedure treats the values outside of the equivalence range as the null-hypothesis. For ease of comparison we can plot 1-SGPV (see Figure \ref{fig:TOSTSGPV2}) to make the values more easily comparable. We see that the *p*-value from the TOST procedure and the SGPV follow each other closely. When we discuss the relationship between the p-values from TOST and the SGPV, we focus on their correspondence at three values, namely where the TOST *p* = 0.025 and SGPV is 1, where the TOST *p* = 0.5 and SGPV = 0.5, and where the TOST *p* = 0.975 and SGPV = 1. These three values are important for the SGPV because they indicate the values at which the SGPV indicates the data should be interpreted as compatible with the null hypothesis (SGPV = 1), or with the alternative hypothesis (SGPV = 0), or when the data are strictly inconclusive (SGPV = 0.5). These three points of overlap are indicated by the horizontal dotted lines in Figure \ref{fig:TOSTSGPV2}.

```{r, TOSTSGPV2, fig.cap="Comparison of *p*-values from TOST (black line) and 1-SGPV (dotted grey line) across a range of observed sample means (x-axis) tested against a mean of 145 in a one-sample *t*-test with a sample size of 30 and a standard deviation of 2."}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 1001),
     yaxt = "n",
     xaxt = "n",
     ylab = "P-value for TOST and 1-SGPV",
     xlab = "Observed Mean")
axis(1, at = seq(0,1000,100), labels = seq(140,150,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)

lines(1-sgpv_list, type="l", col = "darkgrey", lwd = 3, lty = 3)
lines(p_tost_list, lwd = 3)
abline(h = c(0.975, 0.5, 0.025),
       lty = 2,
       col = "grey")
```
```{r, include=FALSE}
m <- 145
mu <- 145
sd <- 2
n <- 30
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.05

@@

Two other points always have to overlap. When the 95% CI falls completely inside the equivalence region, and one endpoint of the confidence interval is exactly equal to one of the equivalence bounds (see situation B in Figure \ref{fig:TOSTSGPV3}) the TOST *p*-value (which relies on a one-sided test) is always 0.025, and the SGPV is 1. The third point where the SGPV and the *p*-value from the TOST procedure should overlap is where the 95% CI falls completely outside of the equivalence range, but one endpoint of the confidence interval is equal to the equivalence bound (see situation C in Figure \ref{fig:TOSTSGPV3}), when the *p*-value will always be 0.975, and the SGPV is 0. Note that this situation is in essence a minimum-effect test [@murphy_statistical_2014]. The goal of a minimum-effect is not just to reject a difference of zero, but to reject the smallest effect size of interest (i.e., the equivalence bounds). An equivalence test and minimum effect test against the same equivalence bound are complementary, and when a TOST *p*-value is larger than 0.975, the *p*-value for the minimum effect test is smaller than 0.05, and the minimum effect test provides no additional information that can not be derived from the *p*-value from the equivalence test. The SGPV summarizes the information from an equivalence test and a minimum-effect test. These can be two relevant questions to ask, although it often makes sense to combine an equivalence test and a null-hypothesis test instead [@lakens_equivalence_2018].

@@

There are three situations where *p*-values for TOST differentiate between observed results, while the SGPV does not differentiate. The first two situations were discussed before and can be seen in Figure \ref{fig:TOSTSGPV1}. When the SGPV is either 0 or 1, *p*-values from the equivalence test fall between 0.975 and 1 or between 0 and 0.025. While the SGPV is 1 as long as the confidence interval falls completely within the equivalence bounds, the *p*-value for the TOST continues to differentiate between results as a function of how far the confidence interval lies within the equivalence bounds (the further the confidence interval is from both bounds, the lower the *p*-value). The easiest way to see this is by plotting the SGPV against the *p*-value from the TOST procedure. The situations where the *p*-values from the TOST procedure continue to differentiate based on how extreme the results are, but the SGPV is a fixed value are indicated by the parts of the curve where there are vertical straight lines at second generation *p*-values of 0 and 1.

>>>>

Second generation *p*-values (SGPV) were recently proposed as a statistic that represents ‘the proportion of data-supported hypotheses that are also null hypotheses’ [@blume_second-generation_2018]. 
The researcher specifies an equivalence range around a null hypothesis of values that are considered practically equivalent to the null hypothesis. 
The SGPV measures the degree to which a set of data-supported parameter values falls within the interval null hypothesis. 
If the estimation interval falls completely within the equivalence range, the SGPV is 1. 
If the confidence interval falls completely outside of the equivalence range, the SGPV is 0. Otherwise the SGPV is a value between 0 and 1 that expresses the overlap of data-supported hypotheses and the equivalence range. 
When calculating the SGPV the set of data-supported parameter values can be represented by a confidence interval (CI), although one could also choose to use credible intervals or Likelihood support intervals (SI). 
When a confidence interval is used, the SGPV and equivalence tests such as the Two One-Sided Tests (TOST) procedure [@lakens_equivalence_2017;@meyners_equivalence_2012;@quertemont_how_2011;@schuirmann_comparison_1987] appear to have close ties, because both tests compare a confidence interval against an equivalence range.
Here, we aim to examine the similarities and differences between the TOST procedure and the SGPV. 
We limit our analysis to continuous data sampled from a bivariate normal distribution.

@@

The second generation *p*-value (SGPV) is calculated as:
$$
  p _ { \delta } = \frac { \left| I \cap H _ { 0 } \right| } { | I | } \times \max \left\{ \frac { | I | } { 2 \left| H _ { 0 } \right| } , 1 \right\}
$$
where I is the interval based on the data (e.g., a 95% confidence interval) and H~0~ is the equivalence range. The first term of this formula implies that the second generation *p*-value is the width of the confidence interval that overlaps with the equivalence range, divided by the total width of the confidence interval. The second term is a 'small sample correction' (which will be discussed later) that comes into play whenever the confidence interval is more than twice as wide as the equivalence range.
To examine the relation between the TOST *p*-value and the SGPV we can calculate both statistics across a range of observed effect sizes. Replicating the example by @blume_second-generation_2018, in Figure \ref{fig:TOSTSGPV1} *p*-values are plotted for the TOST procedure and the SGPV. The statistics are calculated for hypothetical one-sample *t*-tests for observed means ranging from 140 to 150 (on the x-axis). The equivalence range is set to 145 ± 2 (i.e., an equivalence range from 143 to 147), the observed standard deviation is assumed to be 2, and the sample size is 30. For example, for the left-most point in Figure \ref{fig:TOSTSGPV1} the SGPV and the TOST *p*-value is calculated for a hypothetical study with a sample size of 30, an observed standard deviation of 2, and an observed mean of 140, where the *p*-value for the equivalence test is 1, and the SGPV is 0.

@@

```{r, TOSTSGPV1, fig.cap="Comparison of *p*-values from TOST (black line) and SGPV (grey line) across a range of observed sample means (x-axis) tested against a mean of 145 in a one-sample *t*-test with a sample size of 30 and a standard deviation of 2, illustrating that when the TOST *p*-value = 0.5, the SGPV = 0.5, when the TOST *p*-value is 0.975, 1-SGPV = 1, and when the TOST *p*-value = 0.025, 1-SGPV = 0."}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 1001),
     yaxt = "n",
     xaxt = "n",
     ylab = "P-value for TOST and SGPV",
     xlab = "Observed Mean")
axis(1, at = seq(0,1000,100), labels = seq(140,150,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)
lines(sgpv_list, type="l", col = "darkgrey", lwd = 3)
lines(p_tost_list, lwd = 3)
```

Our conclusions about the relationship between TOST *p*-values and SGPV hold for second generation *p*-values calculated from confidence intervals, and assuming data is sampled from a bivariate normal distribution. Readers can explore the relationship between TOST *p*-values and SGPV for themselves in an online Shiny app: http://shiny.ieis.tue.nl/TOST_vs_SGPV/.

The SGPV treats the equivalence range as the null-hypothesis, while the TOST procedure treats the values outside of the equivalence range as the null-hypothesis. For ease of comparison we can plot 1-SGPV (see Figure \ref{fig:TOSTSGPV2}) to make the values more easily comparable. We see that the *p*-value from the TOST procedure and the SGPV follow each other closely. When we discuss the relationship between the *p*-values from TOST and the SGPV, we focus on their correspondence at three values, namely where the TOST *p* = 0.025 and SGPV is 1, where the TOST *p* = 0.5 and SGPV = 0.5, and where the TOST *p* = 0.975 and SGPV = 1. These three values are important for the SGPV because they indicate the values at which the SGPV indicates the data should be interpreted as compatible with the null hypothesis (SGPV = 1), or with the alternative hypothesis (SGPV = 0), or when the data are strictly inconclusive (SGPV = 0.5). These three points of overlap are indicated by the horizontal dotted lines in Figure \ref{fig:TOSTSGPV2} at TOST *p*-values of 0.975, 0.5, and 0.025.

```{r, TOSTSGPV2, fig.cap="Comparison of *p*-values from TOST (black line) and 1-SGPV (grey line) across a range of observed sample means (x-axis) tested against a mean of 145 in a one-sample *t*-test with a sample size of 30 and a standard deviation of 2."}
plot(NA, 
     ylim = c(0, 1), 
     xlim = c(0, 1001),
     yaxt = "n",
     xaxt = "n",
     ylab = "P-value for TOST and 1-SGPV",
     xlab = "Observed Mean")
axis(1, at = seq(0,1000,100), labels = seq(140,150,1), las = 1)
axis(2, at = seq(0,1,0.1), labels = seq(0,1,0.1), las = 1)

lines(1-sgpv_list, type="l", col = "darkgrey", lwd = 3)
lines(p_tost_list, lwd = 3)
abline(h = c(0.975, 0.5, 0.025),
       lty = 2,
       col = "grey")
```
```{r, include=FALSE}
m <- 145
mu <- 145
sd <- 2
n <- 30
low_eqbound = -2 
high_eqbound = 2 
alpha = 0.05

@@

Two other points always have to overlap. When the 95% CI falls completely inside the equivalence region, and one endpoint of the confidence interval is exactly equal to one of the equivalence bounds (see situation B in Figure \ref{fig:TOSTSGPV3}) the TOST *p*-value (which relies on a one-sided test) is always 0.025, and the SGPV is 1. Note that when sample sizes are small or equivalence bounds are narrow, small *p*-values for the TOST or a SGPV = 1 might not be observed in practice if too few observations are collected. The third point where the SGPV and the *p*-value from the TOST procedure should overlap is where the 95% CI falls completely outside of the equivalence range, but one endpoint of the confidence interval is equal to the equivalence bound (see situation C in Figure \ref{fig:TOSTSGPV3}), when the *p*-value will always be 0.975, and the SGPV is 0. Note that this situation is in essence a minimum-effect test [@murphy_statistical_2014]. The goal of a minimum-effect is not just to reject a difference of zero, but to reject the smallest effect size of interest (i.e., the equivalence bounds). An equivalence test and minimum effect test against the same equivalence bound are complementary, and when a TOST *p*-value is larger than 0.975, the *p*-value for the minimum effect test is smaller than 0.05 (and therefore the minimum effect test provides no additional information that can not be derived from the *p*-value from the equivalence test). The SGPV summarizes the information from an equivalence test (and the complementary minimum-effect test). These can be two relevant questions to ask, although it often makes sense to combine an equivalence test and a null-hypothesis test instead [@lakens_equivalence_2018].

@@

There are three situations where *p*-values for TOST differentiate between observed results, while the SGPV does not differentiate. The first two situations were discussed before and can be seen in Figure \ref{fig:TOSTSGPV1}. When the SGPV is either 0 or 1, *p*-values from the equivalence test fall between 0.975 and 1 or between 0 and 0.025. Where the SGPV is 1 as long as the confidence interval falls completely within the equivalence bounds, the *p*-value for the TOST continues to differentiate between results as a function of how far the confidence interval lies within the equivalence bounds (the further the confidence interval is from both bounds, the lower the *p*-value). The easiest way to see this is by plotting the SGPV against the *p*-value from the TOST procedure. The situations where the *p*-values from the TOST procedure continue to differentiate based on how extreme the results are, but the SGPV is a fixed value are indicated by the parts of the curve where there are vertical straight lines at second generation *p*-values of 0 and 1.
